Polka-dotted Stars: a Hierarchical Model for Mapping Stellar Surfaces Using Occultation Light Curves and the Case of TOI-3884

### 2025-02-28 src/tex/ms.tex
wrote up toi3884 results

\begin{equation}
    \label{eq:ptemcee}
    P_{\rm exchange} = \min\left(1, \exp\left[\left(\frac{1}{T_i} - \frac{1}{T_j}\right)\right.\right.\\
    \left.\left.\left(\ln p(\theta_j|\mathcal{D}) - \ln p(\theta_i|\mathcal{D})\right)\right]\right)
\end{equation}

@@

Our analysis of the spot latitude distribution is illustrated in Figure \ref{fig:spot_latitude}. This figure shows the posterior 
probability distribution of spot latitudes derived from our MCMC analysis. The black line represents the mean distribution, 
while the individual pink lines show a subset of 1000 posterior samples, highlighting the inherent variability in our constraints. 
The distribution exhibits two prominent peaks centered at approximately $-75^{\circ}$ and $+75^{\circ}$ latitude, 
with negligible probability near the equator.

@@

The stellar inclination angle was found to be $i_\star = {34.8}^{+5.62}{-6.17}$ degrees, indicating that the stellar rotation 
axis is considerably inclined from the line of sight. More notably, the sky-projected spin-orbit angle (stellar obliquity) was 
determined to be $\lambda\star = {80.37}^{+49.6}_{-27.5}$ degrees.
The stellar rotation period was constrained to $P_\text{rot} = 9.07^{+0.45}_{-0.51}$ days, which places TOI-3884 among the 
more rapidly rotating M dwarfs, consistent with its observed activity level and estimated age.

@@

\bibliography{bib}

>>>>

\begin{equation}
    \label{eq:ptemcee}
    P_{\rm exchange} = \min\left(1, \exp\left[\left(\frac{1}{T_i} - \frac{1}{T_j}\right)\right.\right.\left.\left.\left(\ln p(\theta_j|\mathcal{D}) - \ln p(\theta_i|\mathcal{D})\right)\right]\right)
\end{equation}

@@

Our analysis of the spot latitude distribution is illustrated in Figure \ref{fig:spot_latitudes}. This figure shows the posterior 
probability distribution of spot latitudes derived from the model. The black line represents the mean distribution, 
while the individual pink lines show a subset of 1000 posterior samples, highlighting the inherent variability in our constraints. 
The distribution exhibits two prominent peaks centered at approximately $-75^{\circ}$ and $+75^{\circ}$ latitude, 
with negligible probability near the equator.

@@

The stellar inclination angle was found to be $i_\star = {34.8}^{+5.62}_{-6.17}$ degrees, indicating that the stellar rotation 
axis is considerably inclined from the line of sight. More notably, the sky-projected spin-orbit angle (stellar obliquity) was 
determined to be $\lambda\star = {80.37}^{+49.6}_{-27.5}$ degrees.
The stellar rotation period was constrained to $P_\text{rot} = 9.07^{+0.45}_{-0.51}$ days, which places TOI-3884 among the 
more rapidly rotating M dwarfs, consistent with its observed activity level and estimated age.

@@

Figure \ref{fig:toi3884-moll} presents a visualization of the stellar surface of TOI-3884 with the modeled spot 
distributions from nine different posterior samples. Each panel displays a Mollweide projection of the entire stellar surface, 
with the color scale indicating intensity variations. The dark purple regions represent cooler areas corresponding to stellar 
spots, while the yellow-orange regions depict the warmer, unspotted photosphere.
These visualizations clearly demonstrate the high-latitude spot configurations. Across the different posterior samples, 
we consistently observe spot concentrations near the poles, though with some variation in the exact morphology and distribution. 
The variations between panels illustrate the range of possible spot configurations that are consistent with our transit 
observations. Despite these differences, the predominance of high-latitude spots is a robust feature across all samples, 
supporting the conclusion that TOI-3884 hosts significant near-polar magnetic activity. 
The consistency of these high-latitude features across different posterior samples demonstrates that this result is 
not sensitive to the specific parameter values within our model's posterior distribution.
These surface maps provide a direct visualization of the spot configuration that gives rise to the transit light curve 
asymmetries shown in Figure \ref{fig:toi3884-transits}, illustrating the three-dimensional 
geometry of the TOI-3884 system.

Similarly, Figure \ref{fig:toi3884-traj} shows nine representative views from our posterior samples. Each panel displays the stellar disk as it 
appears from Earth, with the temperature variation across the surface indicated by the color gradient. The black dots trace 
the path of TOI-3884b across the stellar disk during transit, clearly illustrating how the planet's trajectory intersects 
with surface features. As the planet crosses regions of different brightness on the stellar surface, it blocks varying amounts 
of light, creating the subtle but distinctive signatures we observed in the data. The consistent presence of near-polar spots 
across different posterior samples, combined with the specific transit geometry, provides strong evidence for our inferred 
low stellar inclination and high stellar obliquity. These results illustrate the three-dimensional relationship 
between the planet's orbital plane and the star's rotation axis, supporting the significant spin-orbit misalignment in the 
TOI-3884 system.
%
\begin{figure*}[hbt!]
    \centering
    \includegraphics[width=\textwidth]{figures/toi3884-samples-moll.pdf}
    \caption{Mollweide projections of the stellar surface of TOI-3884 showing spot distributions from nine representative 
    posterior samples. The majority of samples exhibit prominent high-latitude spot concentrations, with some variation in 
    the exact morphology and distribution. This consistent pattern of near-polar spots across different posterior samples 
    provides strong support for our inferred high-latitude magnetic activity. 
    These surface configurations produce the transit light curve modulations observed in Figure~\ref{fig:toi3884-transits}.}
    \label{fig:toi3884-moll}
\end{figure*}
%

%
\begin{figure*}[hbt!]
    \centering
    \includegraphics[width=\textwidth]{figures/toi3884-samples-traj.pdf}
    \caption{Observer's view of TOI-3884 during transit events from nine representative posterior samples. 
    Each panel shows the stellar disk as viewed from Earth, with dark purple regions indicating cooler spots and 
    yellow-orange areas representing the warmer photosphere. Black dots trace the path of TOI-3884b across the stellar disk 
    during transit. The consistent presence of high-latitude spots intersecting or near the planet's trajectory explains the 
    asymmetries observed in the transit light curves. This visualization demonstrates how the combination of stellar inclination 
    ($i_\star = {34.8}^{+5.62}_{-6.17}$ degrees), spot distribution, and planetary orbital geometry produces the distinctive
    transit signatures in the TESS data.}
    \label{fig:toi3884-traj}
\end{figure*}
%
\bibliography{bib}
